### 2018-05-07 1_intro.tex
Add some comment ideas. Cleanup a bit.

Current work aims to rework the proof-of-concept tool \porthos{} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and maintainability.
%interprocedural analysis
As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name -- \porthos[2]%
\footnote{Hereinafter with the name `\porthos' we refer to the tool \porthos{} version 1 (also addressed as \porthos[1]), whereas the new version of \porthos{} is called \porthos[2].} %
. Considering the enhancements of the architecture, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.

>>>>

% PROBLEM STATEMENT: MORE CLEAR: 
% 1) Definition of research scope and goals 
% 2) "- References throw light on the topic from a variety of perspectives"
%    "-The thesis combines the cited  works and empirical data consistently and clearly"
%    "- Thesis demonstrates critIcal evaluation of existing knowledge"
% 3) Research Methods
%    "- Appropriate methods have been used in a well-founded manner. Reliability and validity or  trustworthiness of the study/results are evaluated."
% 4) Results and contribution 
%    results + justification
%    Goals and results must match
%    the results must be based on independent research , not on the references.
%    "significance for the field of research or industry in question, and contribution to knowledge"
%    "Cited works must be evaluated critically"
%    "Conclusions SHOULD NOT BE drawn JUST appropriately from the material"

Current work aims to rework the proof-of-concept tool \porthos{} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, extensibility, reliability and maintainability.
%interprocedural analysis
As the general architecture and almost all components of \porthos{} have been redesigned, the tool received a new name -- \porthos[2]%
\footnote{Hereinafter with the name `\porthos' we refer to the tool \porthos{} version 1 (also addressed as \porthos[1]), whereas the new version of \porthos{} is called \porthos[2].} %
. Considering the enhancements of the architecture, \porthos[2] represents a generalised framework for SMT-based memory model-aware analysis, which can not only perform the portability analysis, but can serve as a basis for other kinds of static code analysis.

### 2018-05-07 4_impl.tex
Add some comment ideas. Cleanup a bit.

On the \texttt{Y}-level of abstraction, we define the \texttt{YType} as an alias for a type (since the \ytree{} is not typed, all expressions do not have type, however, the \texttt{YType} is used for storing the information of the type for declarations

@@

The \ytree{} is compiled into the low-level event-based program representation \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF]. The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \cup \xgraph[DF]$.
Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.

@@

  \1 \texttt{XComputationEvent} \\
    We distinguish two types of computation events:
      %\begin{itemize}[noitemsep]
      \2 \texttt{XUnaryComputationEvent} that encodes bit negation and no-operation, 
      \2 \texttt{XBinaryComputationEvent} that encodes numeric operations (such as addition, multiplication, etc.), bit vector operations (such as bit-and, bit-xor, etc.), relative operations (such as greater-then comparison, equality comparison, etc.), and logical operations (such as conjunction and disjunction). 
      %\end{itemize}
    \newline
    The computation event class implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
    This is a model-level optimisation, which is possible because a computation event performs computation over local-only memory and does not change value of any memory unit.
    Thus, the \textit{computation} abstraction (as the CPU time spent for the computation itself) can be safely removed from the model, and the computation event can be seen as a zero-time operation that produces the \textit{value}.
    For analysing large expressions this optimisation is sensible because it considers the whole expression as a single computation event, encoded therefore as a single SMT-variable.
    \newline
    Note, for the purpose of simplifying the X abstraction level, computation events may be modelled as a bodiless function invocation (for instance, `\lstinline{x + y}' is the invocation of the function `\lstinline{+}' to argument `\lstinline{x}' and `\lstinline{y}').

@@

\paragraph{Edges}
Edges are stored as an immutable hash-map from source to target event.
The type of edge is stored in the enumeration \texttt{XEdgeKind}:%TODO: IMPLEMENT THIS
\begin{itemize}[noitemsep]
  \item control-flow edges: a \textit{primary} edge (epsilon or if-then transition) and an \textit{alternative} edge (if-else transition), see more in Section~\ref{ch:port:enc:cf};
  \item data-flow edges: the \co-relation edge and the \rf-relation edge. %TODO! implement and then describe more!
\end{itemize}

@@

@@

The precompiler traverses the \ytree{} and collects information necessary for compiling it into an \xgraph{}.

@@

The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions. % according the rules of propagation. For example, 
Being carried at Y and X representation levels, the type is converted to a Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).
Currently, \porthos[2] operates only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
Array dimension is stored as an integer number at the Y-level (at the X- and Z- levels, \textbf{TODO: think about it}). %the array operations are converted to ).
In addition, \porthos[2] supports ... int\_t ....

   %TODO: support this.

%TODO: describe the XType
%TODO: definitely get rid of bit32

@@

%TODO: <not implemented>
The type analysis includes the function resolution: for correct resolving the semantics of functions, the compiler needs a mapping of each function invocation to its signature.
However due to lack of polymorphism and function overloading mechanism inherent to OOP languages the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis preprocessing unit may be used for analysing the C++ code if needed.

- X-type construction: code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+

- The typification algorithm: propagation rules for each of Y

- size\_t. ...%https://en.wikipedia.org/wiki/C_data_types#stddef.h

@@

As it was discussed in Section~\ref{ch:impl:model:ytree}, the \ytree{} encodes all named objects as \texttt{YVariableRef}.

@@

mapping

@@

- hierarchy of Compilers (XCompiler is an stateful abstract machine)

@@

@@

>>>>

On the \texttt{Y}-level of abstraction, we define the \texttt{YType} as a \textit{reference} for the type (since the \ytree{} is not typed, all expressions do not have type, however, the \texttt{YType} is used for storing the information on declaration, including the type itself, type modifiers and qualifiers).

@@

The \ytree{} is compiled into the low-level event-based program representation \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF]. The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \cup \xgraph[DF]$.
\ Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.

@@

  \1 \texttt{XComputationEvent} \\
    We distinguish two types of computation events:
      %\begin{itemize}[noitemsep]
      \2 \texttt{XUnaryComputationEvent} that encodes bit negation and no-operation, 
      \2 \texttt{XBinaryComputationEvent} that encodes numeric operations (such as addition, multiplication, etc.), bit vector operations (such as bit-and, bit-xor, etc.), relative operations (such as greater-then comparison, equality comparison, etc.), and logical operations (such as conjunction and disjunction). 
      %\end{itemize}
    \newline
    The computation event class implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
    This is a model-level optimisation, which is possible because a computation event performs computation over local-only memory and does not change value of any memory unit.
    Thus, the \textit{computation} abstraction (as the CPU time spent for the computation itself) can be safely removed from the model, and the computation event can be seen as a zero-time operation that produces the \textit{value}.
    For analysing large expressions this optimisation is sensible because it considers the whole expression as a single computation event, encoded therefore as a single SMT-variable.
    \newline
    Note, for the purpose of simplifying the X abstraction level, computation events may be modelled as a bodiless function invocation (for instance, `\lstinline{x + y}' is the invocation of the function `\lstinline{+}' to the arguments `\lstinline{x}' and `\lstinline{y}').

@@

\paragraph{Edges}
Edges are stored as an immutable hash-map from source to target event.
The type of edge is stored in the enumeration \texttt{XEdgeKind}:%TODO: IMPLEMENT THIS
\begin{itemize}[noitemsep]
  \item the \textit{control-flow edges}: a primary edge (epsilon or if-then transition) and an alternative edge (if-else transition), see more in Section~\ref{ch:port:enc:cf};
  \item the \textit{data-flow edges}: the \co-relation edge and the \rf-relation edge. %TODO! implement and then describe more!
\end{itemize}

@@

\paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
%TODO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'
todo

@@

The precompiler traverses the \ytree{} and collects information necessary for its compilation into an \xgraph{}.

@@

The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions. % according the rules of propagation. For example, 
Being carried at Y and X representation levels, the type is converted to a Z-type at the stage of the Z-formula encoding (see Section~\ref{ch:impl:proc:z-encoder}).

%TODO: YOU REALLY NEED TO THINK MORE ABOUT TYPES ON Y, X, Z LEVELS + reconsider this paragraph (and others about types)
Currently, \porthos[2] operates only the primitive C types (such as \texttt{int}, \texttt{char}, \texttt{float}, etc.), which is modelled as an enumeration \texttt{XType}. %TODO: model it.
Array dimension is stored as an integer number at the Y-level (at the X- and Z- levels, \textbf{TODO: think about it}). %the array operations are converted to ).
In addition, \porthos[2] has a built-in extensible database storing the semantics of non-primitive C types stored by the \texttt{X2ZTypeConverter}, %TODO: Or XTypeHookManager?
which can be requested for converting an X-type to a Z-type. %TODO: implement this.
%TODO: describe the XType (before) + implement this clearly! + definitely get rid of bit32

@@

%TODO: <not implemented>
The type analysis includes the process of resolving semantics of function invocations.
The result is stored in a map of function invocation expression to its resolved signature, so that the compiler can decide either to jump to the function body and interpret it, or to hook the invocation if the semantics of the function is resolved.
However, due to lack of polymorphism and function overloading mechanism inherent to OOP languages, the function resolution algorithm may set up the mapping only for the function name and thus may neglect analysis of types of the arguments, we build the mapping for the full function signature so that the type analysis preprocessing unit may be used for analysing the C++ code if needed.

==

The X-type construction: //code of class YType2TypeConverter (todo: rename: it's not conversion, it's a type construction, from the y-level name+qualifiers+
- on Z level, type is one-pointer-level
- The typification algorithm: propagation rules for each of Y
- as vector array as needed.

Y: quantifiers+specifiers+typeName
X: ?.. Z? TODO: rename+rework XType->XZType
Z: Int, Real, Boolean, BitVector[N]

@@

On the compilation stage, once the compiler meets the reference to a variable, it should know whether it is local or global variable.
In general case,

@@

- EXPORT\_SYMBOL "exports the symbol for dynamic linking"

-EXPORT\_SYMBOL\_GPL %https://lwn.net/Articles/154602/

 %https://stackoverflow.com/questions/6670589/how-to-prevent-error-symbol-undeclared-here-despite-export-symbol-in-a-linu

@@

@@

converting recursive structure to plain

stacks: context, readyContext, call, 
  BlockContext (rename): what stores, enum of states

almostReadyContexts: for computation events. Flushed by invocation . When are filled.

setting up edges depending of type of ready-context

stateful. Managers, [graph] builders
  interfaces of XMemoryManager: listing, paragraph

- hierarchy of Compilers: pic on inheritance (XCompiler is an stateful abstract machine). Restrictions of lude-compilers
  postlude: assertion: a new event XAssertionEvent

interfaces that offer compilers: listing of interface

algorithm for interpreting

@@

mapping - as a result

%TODO: kernel's : ACCESS_ONCE() access may be modeled as a volatile memory_order_relaxed access.  %http://www.open-std.org/jtc1/sc22/wg21/docs/papers/2015/n4374.html

% + RCU %https://www.kernel.org/doc/Documentation/RCU/whatisRCU.txt
